#Machine Learning Project

## Overview
The following investigation seeks to predict the manner in which they performed the exercise from other variables in the dataset.

##Data Processing

The data is read in and split into different sets.

```{r}
rm(list =ls())
library(randomForest);library(e1071); library(caret); library(rpart);library(rpart.plot);library(RColorBrewer);library(rattle);
set.seed(3333)

training <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
testing <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
incl_Train <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
Train_1 <- training[incl_Train, ]; Test_1 <- training[-incl_Train, ]
dim(Train_1); dim(Test_1)
```

##Method for Prediction

Attempt 2 methods for prediction: 1)Decision Tree and 2)Random Forest.
 
```{r}
data_NZV <- nearZeroVar(Train_1, saveMetrics=TRUE)
vars_NZV <- names(Train_1) %in% c("new_window", "kurtosis_roll_belt", "kurtosis_picth_belt",
                                  "kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt",
                                  "max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "avg_roll_arm", "stddev_roll_arm",
                                  "var_roll_arm", "avg_pitch_arm", "stddev_pitch_arm", "var_pitch_arm", "avg_yaw_arm",
                                  "stddev_yaw_arm", "var_yaw_arm", "kurtosis_roll_arm", "kurtosis_picth_arm",
                                  "kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm",
                                  "max_roll_arm", "min_roll_arm", "min_pitch_arm", "amplitude_roll_arm", "amplitude_pitch_arm",
                                  "kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell",
                                  "skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell",
                                  "amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm",
                                  "skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_roll_forearm",
                                  "max_yaw_forearm", "min_roll_forearm", "min_yaw_forearm", "amplitude_roll_forearm",
                                  "amplitude_yaw_forearm", "avg_roll_forearm", "stddev_roll_forearm", "var_roll_forearm",
                                  "avg_pitch_forearm", "stddev_pitch_forearm", "var_pitch_forearm", "avg_yaw_forearm",
                                  "stddev_yaw_forearm", "var_yaw_forearm")
Train_1 <- Train_1[!vars_NZV]
dim(Train_1)
Train_1 <- Train_1[c(-1)]

trainingV3 <- Train_1 
for(i in 1:length(Train_1)) { 
  if( sum( is.na( Train_1[, i] ) ) /nrow(Train_1) >= .6 ) { 
    for(j in 1:length(trainingV3)) {
      if( length( grep(names(Train_1[i]), names(trainingV3)[j]) ) ==1)  { 
        trainingV3 <- trainingV3[ , -j] }}}}

dim(trainingV3)
Train_1 <- trainingV3
rm(trainingV3)
col1 <- colnames(Train_1)
col2 <- colnames(Train_1[, -58]) 
Test_1 <- Test_1[col1]
testing <- testing[col2]

dim(Test_1)
dim(testing)
for (i in 1:length(testing) ) {
  for(j in 1:length(Train_1)) {
    if( length( grep(names(Train_1[i]), names(testing)[j]) ) ==1)  {
      class(testing[j]) <- class(Train_1[i])  } } }

testing <- rbind(Train_1[2, -58] , testing) 
testing <- testing[-1,]
modelFit_A <- rpart(classe ~ ., data=Train_1, method="class")
pred_A <- predict(modelFit_A, Test_1, type = "class")
confusionMatrix(pred_A, Test_1$classe)
modelFit_B <- randomForest(classe ~. , data=Train_1)
pred_B <- predict(modelFit_B, Test_1, type = "class")
confusionMatrix(pred_B, Test_1$classe)
pred_test <- predict(modelFit_B, testing, type = "class")
```


##Predict the Testing Set

The following script runs the predictions for a set of data of the same format. In this case, the 20 observations in the test set will be evaluated and answers are exported as text files.

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE) }}

pml_write_files(pred_test)
```